---
title: "Capstone for CRUG"
output: slidy_presentation
---
---


## Major Issue in Using Unsupervised Learning


### Assess the Clustering Tendency


```{r,warning=FALSE, message=FALSE}
library(MASS)
library(fitdistrplus)
library(ggplot2)
library(mclust)
library(cluster)
library(clustertend)
```


### Data preparation

```{r}
f.t.debt.ratio = read.csv(file="C:/Temp/Debt/External_Total_Data.csv", header=TRUE, sep=",") 
#head(f.t.debt.ratio)
# apply log on each columns to normlize data
f.t.df = as.data.frame(round(apply(f.t.debt.ratio, 2, log), 2))

set.seed(123)
hopkins(f.t.df, n = nrow(f.t.df)-1)
```

---

### An illustration of the dataset on two dimension 


```{r}

ggplot(f.t.df, aes(x=F.Debt.Ratio , y=T.Debt.Ratio)) + geom_point() +geom_density2d() 
```


---

## Choose right model

```{r}
f.t.clust  = Mclust(f.t.df)
plot(f.t.clust)
f.t.clust$G
f.t.clust.num = f.t.clust$classification
#write.csv(f.t.clust.num, file="C:/Temp/Debt/log.f.t.clust.num.csv")
```


---

## Check distribution

###Check distribution for each clusted group

1.  Check distribution for cluster 1 group

### Check distribution for external debt ratio
```{r}
clusted.f.t.df = read.csv(file="C:/Temp/Debt/Clusted.f.t.df.csv", header=TRUE, sep=",")
#head(clusted.f.t.df)
attach(clusted.f.t.df)
#Check on external Debt ratio of cluster 1
descdist(F.Debt.Ratio[1:75], discrete = FALSE, boot=1000)

# looks like distribuion is close to normal
#Check normal- accept normal 
norm.fit = fitdist(F.Debt.Ratio[1:75], "norm")
ks.norm= ks.test(F.Debt.Ratio[1:75], "pnorm", mean=norm.fit$estimate[1], sd=norm.fit$estimate[2])
c(ks.norm$statistic, P.Value=ks.norm$p.value)

```


---

### Check distribution for total debt ratio
```{r}
#Check on total debt ratio of cluster 1
descdist(T.Debt.Ratio[1:75], discrete = FALSE, boot=1000)

# looks like distribuion is close to normal
#Check normal- accept normal 
norm.fit = fitdist(T.Debt.Ratio[1:75], "norm")
ks.norm= ks.test(T.Debt.Ratio[1:75], "pnorm", mean=norm.fit$estimate[1], sd=norm.fit$estimate[2])
c(ks.norm$statistic, P.Value=ks.norm$p.value)
```


---


## Check relationship between external debt ratio and total debt ratio


### Check relationshop between external debt ratio and total debt ratio in cluster 1 group


```{r}
x1 = F.Debt.Ratio[1:75]
y1 = T.Debt.Ratio[1:75]
z1 = T.Debt.Ratio[1:75]-F.Debt.Ratio[1:75]
plot(x1, y1, xlab="External Debt Ratio", ylab ="Total Debt Ratio", col='blue')
#check rank
plot(rank(x1), rank(y1)) 
```

---

2. Check distributon on cluster 2 group 

```{r}
#Check on external Debt ratio of cluster 2
descdist(F.Debt.Ratio[76:88], discrete = FALSE, boot=1000)

# looks like distribuion is close to uniform or normal
ks.test(F.Debt.Ratio[76:88],"punif",-1,1) #uniform is rejected

#Check normal- accept normal 
norm.fit = fitdist(F.Debt.Ratio[76:88], "norm")
ks.norm= ks.test(F.Debt.Ratio[76:88], "pnorm", mean=norm.fit$estimate[1], sd=norm.fit$estimate[2])
c(ks.norm$statistic, P.Value=ks.norm$p.value)

#Check on total debt ratio of cluster 2
descdist(T.Debt.Ratio[76:88], discrete = FALSE, boot=1000)

# looks like distribuion is close to uniform or normal
ks.test(T.Debt.Ratio[76:88],"punif",-1,1) #uniform is rejected
#Check normal- accept normal 
norm.fit = fitdist(T.Debt.Ratio[76:88], "norm")
ks.norm= ks.test(T.Debt.Ratio[76:88], "pnorm", mean=norm.fit$estimate[1], sd=norm.fit$estimate[2])
c(ks.norm$statistic, P.Value=ks.norm$p.value)
```

---

###  Check relationshop between external debt ratio and total debt ratio in cluster 2 group

```{r}
x2 = F.Debt.Ratio[76:88]
y2 = T.Debt.Ratio[76:88]
z2 = T.Debt.Ratio[76:88]-F.Debt.Ratio[76:88]
# external and total
plot(x2, y2, xlab="External Debt Ratio", ylab ="Total Debt Ratio", col='blue')
abline(h=0.1,col='red')

#check rank
plot(rank(x2), rank(y2))
```

---
### Further check bewteen external debt ratio and domestic debt ratio

 Plot on cluster 2-External debt ratio and domestic debt ratio

```{r}
#External debt ratio and domestic debt ratio
plot(x2, z2, xlab="External Debt Ratio", ylab ="Domestic Debt Ratio")

```


---
## Capture your finding 

### Timeline plot on two clusters output

```{r, warning=FALSE, message=FALSE}
library(ggplot2, warn.conflicts=FALSE)

time.clust=read.csv(file="C:/Temp/Debt/Time.Clusted.f.t.df.1.csv", header=TRUE, sep=",") 

# Default countries based on timeline
ggplot(time.clust, aes(x=Year, y=ClustNo, color=Country, shape=Region)) + geom_point(size=5) + ggtitle("Cluster Groups Based on Timeline ") + xlab("Year") + ylab("Cluster No.") + ylim (0, 4)  + theme(axis.text=element_text(size= 8, face='bold'), axis.title=element_text(size= 8,face="bold")) + theme(plot.title = element_text(size=8, face='bold')) + theme(legend.text=element_text(size=8, face='bold')) + theme(legend.title = element_text(size = 8, face = 'bold')) 
```


---

### Number of default occurs for each country from two clusers
```{r}
#Frequencey of default countries 
ggplot(time.clust, aes(x=Year, y=ClustNo, color=Country)) + geom_point() + facet_wrap(Region~Country) +ggtitle("Country & Region within Cluster group") + ylim(0,5) + theme(axis.text=element_text(size=12, face='bold'), axis.title=element_text(size=12,face="bold")) 
```

---

### Domestic is overlooked - Domestic debt adjusted by inflation over 10 years

```{r}
library(lattice)
inf = read.csv(file="C:/Temp/Debt/Debt_Inflation.csv", header=TRUE, sep=",") 
#head(inf)
xyplot(debt.adj.inf ~ year.a.d |id, data = inf, type="l", subset=(id < 29), strip=FALSE, xlab=" Ten Years", ylab = "Debt Adjuste Inf", main="Debt Adjusted by Inflation over Ten Years")

```